```{r  echo=FALSE }
opts_knit$set(warning=FALSE, message=FALSE, commment=NA)
# start with clean bib
require(knitcitations)
cleanbib()
````
I am finding myself more and more drawn to markdown rather then tex/Rnw as my standard format (not least of which is the ease of displaying the files on github, particularly now that we have automatic image uploading).  One thing I miss from latex is the citation commands. (I understand these can be provided to markdown via Pandoc, but I'd like to simply have to knit the document, and not then run it through pandoc, latex, or another interpreter).  I've taken a little whack at generating in-text citations using knitr and other R tools. 

### DOI Approach

I've put some simple functions in a `knitcitations` package.  The functions use the crossref API to grab citation information given a doi, so I don't have to generate a bibtex file for papers I'm reading, (inspired by the [kcite](http://wordpress.org/extend/plugins/kcite/) package for Wordpress).  One can grab my package from github
```{r  eval=FALSE }
library(devtools)
install_github("knitcitations", "cboettig")
````

and load the package
```{r  }
require(knitcitations)
````

Then we can generate a citation given a DOI with the `ref` function:

```{r  results="asis"}
ref("10.1111/j.1461-0248.2005.00827.x")
````

We can generate inline citations in the short name/date format with the `citet` function:

```{r}
citet("10.1111/j.1461-0248.2005.00827.x")
````

Similarly we can generate parenthetical citations with the `citep` function, 

```{r}
citep(c("10.1111/j.1461-0248.2005.00827.x","10.1890/11-0011.1"))
````

Which can take a list of DOIs to cite parenthetically.  The `citet` and `citep` functions are automatically retrieving the available metadata via the Crossref API, and R is storing the information to generate the final bibliography.  


### Bibtex Approach

If we have a bibtex file, we can use this for the citations as well.  If you don't have a bibtex file handy, you can make one containing the citation information for some of R's packages: 

```{r }
write.bibtex(c('bibtex', 'knitr', 'knitcitations'), file="example.bib")
````

We could also have used the `ref` function write a bibtex file for our list of DOIs

```{r }
refs <- lapply(c("10.1111/j.1461-0248.2005.00827.x","10.1890/11-0011.1"), ref)
write.bibtex(refs, file="refs.bib")
```


Once we have a bibtex file available, we must read it into R.   

```{r  bibtex, results="asis"}
bib <- read.bibtex("example.bib")
````

We can now create citations from `bib` using the bibtex key,

```{r results="asis"}
bib[["knitr"]]
````

The inline citation tools can also now use this `bib` instead of a DOI to generate a citation,

```{r }
citet(bib[["knitr"]])
citep(bib[c("knitr", "bibtex")] )
````


### Using the inline citations 

The inline citation calls are designed to be used with knitr's inline code blocks.  In markdown, these are enclosed in \`r \`.  The output format will use the plain-text rendering rather than the code markup.  Thus we can use the line `citep("10.1111/j.1461-0248.2005.00827.x")` to generate a parenthetical citation, `r citep("10.1111/j.1461-0248.2005.00827.x")`. We can generate the in-text citations with `citet`, such as `r citet(bib[["knitr"]])`.  

### Generating the final bibliography
As we go along adding inline citations, R stores the list of citation info.  Then at the end of the document, use this command to print the bibliography generated by the use of our inline citations. 

```{r  results="asis"}
bibliography()
````

I hope to add markup to format this a bit more nicely later.  For instance, we want the links to appear as real links.  Additionally, we may want to add markup around the citations, such as the reason for the citation into the link using the [Citation Typing Ontology](http://speroni.web.cs.unibo.it/cgi-bin/lode/req.py?req=http:/purl.org/spar/cito). 






