knitcitations
=============

- **Author**: [Carl Boettiger](http://www.carlboettiger.info/)
- **License**: [CC0](http://creativecommons.org/publicdomain/zero/1.0/)
- [Package source code on Github](https://github.com/cboettig/knitcitations)
- [Online package documentation](http://cboettig.github.com/knitcitations/index.html)


`knitcitations` is an R package designed to add dynamic citations to dynamic documents created with [Yihui's knitr package](https://github.com/yihui/knitr).



Installation 
------------

Install directly from Github using [Hadley's devtools package](https://github.com/hadley/devtools)


You can also clone or download the repository and install with `R CMD INSTALL`. Once I'm through the early development phase a copy will be available on CRAN.


Quick start
-----------

It is usually good to clear the bibliographic environment after loading the library, in case any citations are already stored there.  


```{r load, echo=FALSE}
library(knitcitations)
newbib()
```

Cite an article by DOI and the full citation information is gathered automatically.  

```{r results="asis", comment=NA}
citep("10.1890/11-0011.1")
```

Typically this is done with knitr's inline code syntax, creating a parenthetical citation in the text like this `r citep("10.1890/11-0011.1")`.  We display the command in code blocks only for documenting purposes here.  `citep` provides a parenthetical citation; a in-text citation is generated with `citet`, such as this sentence cites `r citet("10.1890/11-0011.1")`.  

When the citation is called, a key in the format `LastNameYear` is automatically created for this citation, so we can now continue to cite this article without remembering the DOI, using the command:

```{r results="asis"}
citep("Abrams2012")
```

Note that a custom key can also be given by naming the DOI, such as `citep(c(AbramsEtAl="10.1890/11-0011.1"))`.


If we have a bibtex file, we can cite such articles by those keys as well.  To demonstrate, let's first create a bibtex file with the citation information for some R packages, using the `bibtex` package utilities (a dependency of `knitcitations`):

```{r message=FALSE}
write.bibtex(c("Yihui2013" = citation("knitr"), "Boettiger2013" = citation("knitcitations"), "TempleLang2012"=citation("RCurl")))
bib <- read.bibtex("knitcitations.bib")
```

Now we can generate citations from `bib`

```{r message=FALSE}
citep(bib[[2]])
citep(bib["Yihui2013"])
```



Generate the final bibliography in a knitr block using the chunk option `results='asis'` to display as text rather than code:  

```{r results="asis"}
bibliography()
```


Other formats can be given as options to `bibliography`, as described in the help documentation, `?bibliography`.  [See the full tutorial](https://github.com/cboettig/knitcitations/blob/master/inst/examples/citations.md) for more about knitcitations.  

## Semantic RDFa

Knitcitations can also return semantic RDFa markup around your citations.  Simply select the "rdfa" in the bibliography option.  


```{r results="asis"}
bibliography("rdfa")
```


Here's what the source HTML looks like (generated as R output by not setting `results="asis"` in the knitr block):

```{r }
bibliography("rdfa")
```

### CiTO  

Additional semantic markup can be added the the citations themselves, such as the reason for the citation.  For instance, we can identify that we have used the method from `r citet("10.1186/2041-1480-1-S1-S6", cito = "usesMethodIn")` with the inline command `citet("10.1186/2041-1480-1-S1-S6", cito = "usesMethodIn")`.  

